% testo della versione inglese del libretto
%%%
% text of the English version of the booklet

\onelineheader{P}{rocessional hymn}{-0.5mm}

\titleofsong{Vieni dal Libano (Come here from Lebanon)}

\textofsong{\songA}

\newpage

\onelineheader{I}{ntroductory Rites}{-0.5mm}

\Celebrantbeginning{In the name of the Father, and of the Son, \cross{} and of the Holy Spirit}.\\

\Assemblybeginning{Amen.}\\

\Celebrant{The grace of our Lord Jesus Christ, and the love of God,
and the communion of the Holy Spirit be with you all.}\\

\Assembly{And with your spirit.}

\vspace{10 mm}

\onelineheader{M}{emory of the Baptism}{-0.5mm}

\Celebrant{\bride{} and \groom,
the Church shares in your joy
and, together with your loved ones,
welcomes you with great affection
in the day in which you decide to make the communion of all life before God, our Father.\\

In this day of celebration, may the Lord listen to you.
May he send his help from heaven and keep you safe.
May he realize the wishes of your hearts
and grant your prayers.\\

Grateful for becoming sons in the Son,
we now remember the Baptism.
From it, as from a fertile seed, the commitment
to live faithfully in love arises and gets stronger.\\

Father,
in the Baptism of thy Son Jesus at the Jordan River
you revealed to the world your spousal love for your people.}\\

\Assembly{We praise you and we give you thanks.}\\

\Celebrant{Jesus Christ,
from your open wound on the Cross you
generated the Church,
your beloved bride.}\\

\Assembly{We praise you and we give you thanks.}\\

\Celebrant{Holy Spirit,
power of the Father and the Son,
today you make the wedding garment of the Church shine in \bride{} and \groom{}.
}\\

\Assembly{We praise you and we give you thanks.}\\

\Celebrant{Almighty God,
origin and source of life,
you renewed us in water
with the power of your Spirit.
Revive in us all the grace of the Baptism,
and grant to \bride{} and \groom{} a free heart and a fervent faith
so that, with a purified spirit,
they might welcome the gift of Marriage,
new path of their sanctification.
Through Christ our Lord.}\\

\Assembly{Amen.}

\vspace{10mm}

\titleofsong{Glory to God in the highest}

\textofsong{\songBB}

\vspace{10mm}

\noindent\secondaryheader{Collect}

\Celebrant{Be attentive to our prayers, O Lord, and in your kindness
pour out your grace on \bride{} and \groom, that are coming together before your altar; may they
be confirmed in love for one another. Through our Lord Jesus Christ, your Son, who lives and reigns with you in the unity of the Holy Spirit, one God, for ever and ever.}\\

\Assembly{Amen.}

\vspace{10mm}

\onelineheader{L}{iturgy of the word}{0mm}

\reading{A reading from the book of Isaiah}{62,\,1--5}{As a bridegroom
rejoices in his bride, so shall your God rejoice in you.}

\biginitial{F}{or} Zion's sake I will not be silent,
for Jerusalem's sake I will not be quiet, until her
vindication shines forth like the dawn and her victory like a burning torch.
Nations shall behold your vindication, and all kings your glory;
you shall be called by a new name pronounced by the mouth of the Lord.
You shall be a glorious crown in the hand of the Lord, a royal diadem held by your God.
No more shall men call you ``Forsaken,'' or your land ``Desolate,''
but you shall be called ``My Delight,'' and your land ``Espoused.''
For the Lord delights in you, and makes your land his spouse.
As a young man marries a virgin, your Builder shall marry you;
and as a bridegroom rejoices in his bride, so shall your God rejoice in you.\\

The word of the Lord.\\

\Assembly{Thanks be to God.}

\vspace{10mm}

\nameofbook{Responsorial Psalm} {\small(from the Psalm 32)}\\

\Assembly{In the Lord our hearts rejoice.}\\

Blessed is the nation whose God is the Lord,\\
the people chosen as his inheritance.\\
Behold, the eye of the Lord is upon those who fear him,\\
upon those who count on his mercy.\\

\Assembly{In the Lord our hearts rejoice.}\\

Our soul waits for the Lord,\\
he is our help and shield.\\
For in him our hearts rejoice;\\
in his holy name we trust.\\

\Assembly{In the Lord our hearts rejoice.}\\

May your mercy, Lord, be upon us;\\
as we put our hope in you.\\

\Assembly{In the Lord our hearts rejoice.}

\newpage

\reading{A reading from the first letter of St.\ Paul to the Corinthians\quad\,\,}
{12,\,31b -- 14,\,1a}{Love will never end.}

\biginitial{B}{rothers}, I shall show you a still more excellent way.
If I speak in human and angelic tongues but do not have love,
I am a resounding gong or a clashing cymbal.
And if I have the gift of prophecy and comprehend all
mysteries and all knowledge; if I have all faith so
as to move mountains but do not have love, I am nothing.
If I give away everything I own, and if I hand
my body over so that I may boast but do not have love, I gain nothing.
 Love is patient, love is kind. It is not jealous,
 (love) is not pompous, it is not inflated,
 it is not rude, it does not seek its own interests,
 it is not quick-tempered, it does not brood over injury,
 it does not rejoice over wrongdoing but rejoices with the truth.
It bears all things, believes all things, hopes all things, endures all things.
Love never fails. If there are prophecies,
they will be brought to nothing; if tongues, they will cease;
if knowledge, it will be brought to nothing.
When I was a child, I used to talk as a child, think
as a child, reason as a child; when I became a man, I put aside childish things.
At present we see indistinctly, as in a mirror, but
then face to face. At present I know partially;
then I shall know fully, as I am fully known.
So faith, hope, love remain, these three;
but the greatest of these is love. Pursue love.\\

The word of the Lord.\\

\Assembly{Thanks be to God.}
 
\vspace{10mm}

\titleofsong{Alleluja}

\newpage

\gospel{A reading from the holy Gospel of Matthew}{6,\,24--34}{Do not worry about tomorrow}

\biginitial{J}{esus} told his disciples: \guillemotleft
No one can serve two masters. He will either hate one and
love the other, or be devoted to one and despise the other.
You cannot serve God and mammon. Therefore I tell you,
do not worry about your life, what you will eat (or drink),
or about your body, what you will wear. Is not life more
than food and the body more than clothing?
Look at the birds in the sky; they do not sow or reap,
they gather nothing into barns, yet your heavenly Father feeds them.
Are not you more important than they?
Can any of you by worrying add a single moment to your life-span? 
Why are you anxious about clothes? Learn from the way the wild flowers grow.
They do not work or spin.
But I tell you that not even Solomon in all his splendor was clothed like one of them.
If God so clothes the grass of the field, which grows today
and is thrown into the oven tomorrow, will he not much
more provide for you, O you of little faith?
So do not worry and say ``What are we to eat?'' or ``What are we to drink?'' or
``What are we to wear?''
All these things the pagans seek. Your heavenly Father knows that you need them all.
But seek first the kingdom (of God) and his righteousness,
and all these things will be given you besides.
Do not worry about tomorrow; tomorrow will take care of itself.
Sufficient for a day is its own evil.\guillemotright.\\

The Gospel of the Lord.\\

\Assembly{Praise to you, Lord Jesus Christ.}

\vspace{10mm}

\onelineheader{H}{omily}{0mm}

\newpage

\onelineheader{R}{ite of Marriage}{-0.5mm}

\Celebrant{Dear \bride{} and \groom, you have come together
in this church so that the Lord may seal and strengthen
your love in the presence of the Church's minister and this community.
Christ abundantly blesses this love.
He has already consecrated you in the Baptism and now he enriches
and strengthens you by the nuptial sacrament, so that you may
assume the duties of marriage in mutual and lasting fidelity.
And so, in the presence of the Church, I ask you to state your intentions.}

\vspace{10mm}

\brideandgroomsym{At the end of the path of engagement,
enlighted by the Holy Spirit, and
with the support of the Christian community,
we came here freely and without reservation 
in the house of the Father,
so that our love shall receive the seal of consecration.\\

Aware of our decision,
we are willing,
with the grace of God,
to love and support each other
for all the days of our life.\\

We shall accept lovingly the children
that God shall donate us
and we shall bring them up according to the Word of Christ 
and the teachings of the Church.\\

We ask you, brothers and sisters,
to pray with us and for us
so that our family
might spread light, peace and joy in the world.}

\vspace{\baselineskip}

\newpage

\secondaryheader{Consent}

\Celebrant{Since it is your intention to enter into marriage,
join your right hands, and declare your consent before God and his church.}\\

\groomsym{I, \groom, take you, \bride, to be my wife.  
With the grace of God, I promise to be true to you
in good times and in bad,
in sickness and in health.  
I will love you and honor you all the days of my life.}\\

\medskip

\bridesym{I, \bride, take you, \groom, to be my husband.
With the grace of God, I promise to be true to you
in good times and in bad, in sickness and in health.  
I will love you and honor you all the days of my life.}\\

\Celebrant{You have declared your consent before the Church.  
May the Lord in his goodness strengthen your consent
and fill you both with his blessings.
What God has joined, men must not divide.}\\

\Assembly{Amen.}

\vspace{10mm}

\onelineheader{B}{lessing of the rings}{0mm}

\Celebrant{May the Lord bless \cross{} these rings:
those who wear them may always have a deep faith in each other.
May they do your will and always live together
in peace, good will, and love.
We ask this through Christ our Lord.}\\

\Assembly{Amen.}\\

\newpage

\groomsym{\bride, take this ring as a sign of my love and fidelity.  
In the name of the Father, and of the Son, and of the Holy Spirit.}\\

\medskip

\bridesym{\groom, take this ring as a sign of my love and fidelity.  
In the name of the Father, and of the Son, and of the Holy Spirit.}\\

\vspace{5mm}

\twolinesheader{P}{rayers of the faithful}{2mm}{and general intercessions}

\Celebrant{Brothers and sisters,
aware of the peculiar gift of grace and love,
by means of which God has been willing to make the love of our brothers \bride{} and \groom{}
perfect and holy, we pray the Lord that,
helped by the example and intercession of the saints,
they shall keep their marriage in fidelity.\\

Let us pray to the Lord: ``\answerforprayers''}\\

\Assembly{\answerforprayers}\\

For \bride{} and \groom, may the vows of love, that today they exchanged in the Lord,
help them building their family according to the word of the Gospel. Let us pray.\\

\Assembly{\answerforprayers}\\

For \bride{} and \groom, united today in the love of Christ,
may the Lord, who rejoices in their love, fill the days of
their lives with blessings. Let us pray.\\

\Assembly{\answerforprayers}\\

For the parents, the relatives and the friends, may they be always a support for 
\bride{} and \groom, both with their hearts and with their prayers,
wherever their home might be. Let us pray.\\

\Assembly{\answerforprayers}\\

May the Holy Spirit be a guide for those who are betrothed
and may it renew in the married couples the grace of the sacrament
of the marriage. Let us pray.\\

\Assembly{\answerforprayers}\\

For the Church of Christ: may it become more and more a family,
united by the patience of listening,
the richness of giving and the joy of forgiving. Let us pray.\\

\Assembly{\answerforprayers}\\

\Celebrant{In communion with the entire heavenly Church
let us invoke the intercession of the Saints.}

\[\begin{array}{ll}
\textrm{Holy Mary, Mother of God} & \textbf{pray for us}\\
\textrm{Holy Mary, Mother of the Church} & \textbf{pray for us}\\
\textrm{Holy Mary, Queen of the family} & \textbf{pray for us}\\
\textrm{Saint Joseph, Mary's husband} & \textbf{pray for us}\\
\textrm{All you angels of God} & \textbf{pray for us}\\
\textrm{Saints Joachim and Ann} & \textbf{pray for us}\\
\textrm{Saints Zaccaria and Elisabeth} & \textbf{pray for us}\\
\textrm{Saint John the Baptist} & \textbf{pray for us}\\
\textrm{Saints Peter and Paul} & \textbf{pray for us}\\
\textrm{Saints Apostles and Evangelists} & \textbf{pray for us}\\
\textrm{All you Holy Martyrs} & \textbf{pray for us}\\
\textrm{Saints Aquila and Priscilla} & \textbf{pray for us}\\
\end{array}\]
\[\begin{array}{ll}
\textrm{Saints Mario and Marta} & \textbf{pray for us}\\
\textrm{Saint Monica} & \textbf{pray for us}\\
\textrm{Saint Paolino} & \textbf{pray for us}\\
\textrm{Saint Brigida} & \textbf{pray for us}\\
\textrm{Saint Rita} & \textbf{pray for us}\\
\textrm{Saint Francesca Romana} & \textbf{pray for us}\\
\textrm{Saint Thomas More} & \textbf{pray for us}\\
\textrm{Saint Giovanna Beretta Molla} & \textbf{pray for us}\\
\textrm{Saint Francis of Assisi} & \textbf{pray for us}\\
\patronsaintofthebride\\
\patronsaintofthegroom\\
\patronsaintofthechurch\\
\textrm{All you holy Saints} & \textbf{pray for us.}\\
\end{array}\]

\Celebrant{Lord, effuse on \bride{} and \groom{}
the Spirit of your love,
may they become a single heart and a single soul:
nothing shall part the couple that you united,
and full of your blessings may they be impermeable to affliction.
Through Christ our Lord.}\\

\Assembly{Amen.}

\vspace{10mm}

\secondaryheader{Profession of Faith}

\textbf{I believe in one God,\\
the Father almighty,\\
maker of heaven and earth,\\
of all things visible and invisible.\\
I believe in one Lord, Jesus Christ,\\
the only begotten Son of God,\\
born of the Father before all ages,\\
God from God, Light from Light, true God from true God,\\
begotten, not made, consubstantial with the Father.\\
Through him all things were made.\\
For us men and for our salvation he came down from heaven,\\
and by the power of the Holy Spirit\\
was incarnate of the Virgin Mary,\\
and became man.\\
For our sake he was crucified under Pontius Pilate;\\
he suffered death, and was buried,\\
and rose again on the third day in accordance with the Scriptures.\\
He ascended into heaven\\
and is seated at the right hand of the Father.\\
He will come again in glory\\
to judge the living and the dead,\\
and his kingdom will have no end.\\
I believe in the Holy Spirit, the Lord, the giver of life,\\
who proceeds from the Father and the Son,\\
who with the Father and the Son is adored and glorified,\\
who has spoken through the Prophets.\\
I believe in one, holy, catholic and apostolic Church.\\
I confess one baptism for the forgiveness of sins\\
and I look forward to the resurrection of the dead,\\
and the life of the world to come. Amen.}

\vspace{15mm}

\onelineheader{E}{ucaristic liturgy}{0mm}

\titleofsong{Segni del tuo amore (Signs of your love)}

\textofsong{\songC}

\vspace{5mm}

\secondaryheader{Prayer over the gifts}

\Celebrant{Pray, my brothers and sisters, that, bringing
to the altar everyday joy and labour, our sacrifice
may be acceptable to God, the almighty Father.}\\

\Assembly{May the Lord accept the sacrifice at your hands
for the praise and glory of his name, for our good, and the good of all his Church.}\\

\Celebrant{Receive, we pray, O Lord,
the offering made on the occasion
of this sealing of the sacred bond of Marriage,
and, just as your goodness is its origin,
may your providence guide its course.
Through Christ our Lord.}\\

\Assembly{Amen.}

\vspace{10mm}

\secondaryheader{Eucaristic prayer}

\Celebrant{The Lord be with you.}\\

\Assembly{And with your spirit.}\\

\Celebrant{Lift up your hearts.}\\

\Assembly{We lift them up to the Lord.}\\

\Celebrant{Let us give thanks to the Lord our God.}\\

\Assembly{It is right and just}.\\

\Celebrant{It is truly right and just, our duty and our salvation,
always and everywhere to give you thanks,
Lord, holy Father, almighty and eternal God.\\

For you willed that the human race,
created by the gift of your goodness,
should be raised to such high dignity
that in the union of husband and wife
you might bestow a true image of your love.\\

For those you created out of charity,
you call to the law of love without ceasing
and you grant them a share in your eternal communion.\\

And so, the sacrament of holy Matrimony,
as the abiding sign of your own love,
consecrates the love of man and woman,
through Christ our Lord.\\

Through him, with the angels and all the saints,
we sing the hymn of your praise, as without end we acclaim:}\newline\newline

\titleofsong{Holy, Holy, Holy}

\textofsong{\songDD}

\vspace{5mm}

\Celebrant{You are indeed Holy, O Lord, and all you have created
rightly gives you praise, for through your Son our Lord Jesus Christ,
by the power and working of the Holy Spirit, you give life
to all things and make them holy, 
and you never cease to gather a people to yourself,
so that from the rising of the sun to its setting a pure
sacrifice may be offered to your name. \\

Therefore, O Lord, we humbly implore you: by the same Spirit
graciously sanctify these gifts that we have brought to you for
consecration; that they may become the Body
and \cross{} Blood of your Son our Lord Jesus Christ
at whose command we celebrate these mysteries.\\

For on the night he was betrayed he himself took bread,
and giving you thanks he said the blessing, broke the bread
and gave it to his disciples, saying:

\medskip

{\emph{Take this, all of you, and eat of it:\\
for this is my Body\\
which will be given up for you}.\par}

\medskip

In a similar way, when supper was ended, he took the chalice,
and giving you thanks he said the blessing, and gave the chalice to his disciples, saying:

\medskip

{\emph{Take this, all of you, and drink from it:\\
for this is  the chalice of my Blood,\\ 
the blood of the new and eternal covenant;\\
which will be poured out for you and for many\\
for the forgiveness of sins. Do this in memory of me}.\par}

\medskip

The mistery of faith.}\\

\Assembly{We proclaim your death, O Lord,\\
and profess your resurrection,\\
until you come again.}\\

\Celebrant{Therefore, O Lord, we celebrate the memorial
of the saving passion of your Son, his wondrous
resurrection and ascension into heaven, and as we look
forward to his second coming, we offer to you in
thanksgiving this holy and living sacrifice.\\

Look, we pray, upon the oblation of your Church,
and, recognizing the sacrificial victim
by whose death you willed to reconcile us to
yourself, grant that we, who are nourished by
the Body and Blood of your Son and filled with his Holy Spirit,
may become one body and one spirit in Christ.\\

May he make of us an eternal offering to you,
so that we may obtain an inheritance with your elect,
especially with the most blessed Virgin Mary, Mother of God,
with your blessed apostles and glorious martyrs and with all the saints,
on whose constant intercession in your presence we rely for unfailing help.\\

May this sacrifice of our reconciliation, we pray,
O Lord, advance the peace and salvation of all the world.
Be pleased to confirm in faith and charity
your pilgrim Church on earth, with your servant and our Pope \nameofpope,
and our Bishop \nameofbishop,
the Order of Bishops, all the clergy,
and the entire people you have gained for your own.\\

Help your children \bride{} and \groom, that in Christ
have built a new family, small Church and sacrament for your love;
may the grace of this day extend to their entire life.\\

Listen graciously to the prayers of this family, whom you have
summoned before you: in your compassion, O merciful Father,
gather to yourself all your children scattered throughout the world.
To our departed brothers and sisters and to all who were pleasing
to you at their passing from this life, give kind admittance to your kingdom.
There, we hope to enjoy forever the fullness of your glory
through Christ our Lord, through whom you bestow on the world all that is good.\\

Through him, and with him, and in him, O God, almighty Father,
in the unity of the Holy Spirit, all glory and honor is yours, for ever and ever.}\\

\Assembly{Amen.}

\vspace{\baselineskip}
\vspace{4mm}

\onelineheader{L}{iturgy of Communion}{0mm}

\Celebrant{Lord gave us his Spirit.
With the faith and the freedom of sons, let us pray together:}\\

\textbf{Our Father, who art in heaven,\\
hallowed be thy name;\\
thy kingdom come,\\
thy will be done\\
on earth as it is in heaven.\\
Give us this day our daily bread,\\
and forgive us our trespasses,\\
as we forgive those who trespass against us;\\
and lead us not into temptation,\\
but deliver us from evil.}

\vspace{10mm}

\secondaryheader{Nuptial Blessing}

\Celebrant{Holy Father, maker of the whole world,
who created man and woman in your own image
and willed that their union be crowned with your blessing,
we humbly beseech you for these servants of yours, \bride{} and \groom,
who are joined today in the sacrament of Matrimony. We praise you and we bless you.\\

\Assembly{Your love for us is ethernal.}\\

\Celebrant{May your abundant blessing, Lord,
come down upon this bride, \bride{},
and upon \groom{}, her companion for life,
and may the power of your Holy Spirit
set their hearts aflame from on high,
so that, living out together the gift of Matrimony,
they may adorn their family with children and
enrich the Church.
We pray you, O Lord.}\\

\Assembly{Hear our prayer.}\\

\Celebrant{In happiness may they praise you, O Lord,
in sorrow may they seek you out;
may they have the joy of your presence
to assist them in their toil,
and know that you are near
to comfort them in their need;
let them pray to you in the holy assembly
and bear witness to you in the world,
and after a happy old age,
together with the circle of friends that surrounds them,
may they come to the kingdom of heaven.
Through Christ our Lord.}\\

\Assembly{Amen.}

\vspace{8mm}

\secondaryheader{Sign of Peace}

\Celebrant{The peace of the Lord be with you always.}\\

\Assembly{And with your spirit.}\\

\Celebrant{Let us offer each other a sign of peace.}\\

\newpage

\titleofsong{Pace sia, pace a voi (Peace be, peace be with you)}

\textofsong{\songE}

\vspace{10mm}

\secondaryheader{Breaking of the Bread}

\textbf{Lamb of God, you take away the sins of the world: have mercy on us.
Lamb of God, you take away the sins of the world: have mercy on us.
Lamb of God, you take away the sins of the world: grant us peace.}\\

\vspace{4mm}

\Celebrant{This is the Lamb of God who takes away the sins of the world.
Happy are those who are called to his supper.}\\

\Assembly{Lord, I am not worthy that you shall enter under my roof,
but only say a word and my soul shall be healed.}

\newpage

\onelineheader{C}{ommunion song}{0mm}

\vspace{-3mm}

\titleofsong{Isaia 62 (Isaiah 62)}

\vspace{-1mm}

\textofsong{\songF}

\Celebrant{Grant, we pray, almighty God,
that the power of the sacrament we have received
may find growth in these servants of yours
and that the effects of the sacrifice we have offered
may be felt by us all.\\

Through Christ our Lord.}\\

\Assembly{Amen.}

\vspace{10mm}

\onelineheader{C}{oncluding rites}{-0.5mm}

\vspace{-5mm}

\Celebrant{Dearest \bride{} and \groom, you celebrated the sacrament of Matrimony
showing your consent in front of me and your witnesses. 
In addition to the divine grace and the effects established by the sacred laws,
your marriage has also some civil effects, in accordance to the Italian laws. 
I hereby read for your knowledge the articles of the Italian Civil Code which regulate
the rights and duties of the marriage: you are expected to respect
them and to comply with them.

\medskip

Art.\ 143: With the marriage the husband and wife acquire
the same rights and the same duties. From the marriage comes
the mutual duty of loyalty, the moral and material assistance,
the cooperation in the interest of the family and the cohabitation.
Both the partners are responsible, each one in relation
to their means and ability of professional or homemade work,
to contribute to the needs of the family.

\medskip

Art.\ 144: The partners both agree on the family life address
and decide their residence according to the prior needs of both.
Each one has the power to implement the address upon which they agreed.

\medskip

Art.\ 147: The marriage commands to both the partners the duty to support, teach
and educate their children according to their abilities,
their natural predisposition and their aspirations.}

\newpage

\onelineheader{S}{olemn blessing at the end of mass}{0mm}

\Celebrant{May God the Eternal Father
keep you of one heart and in love for one another;
may the peace of Christ may dwell in you
and abide always in your home.}\\

\Assembly{Amen.}\\

\Celebrant{May you be blessed in your children,
have solace in your friends
and enjoy true peace with everyone.}\\

\Assembly{Amen.}\\

\Celebrant{May you be witnesses in the world to God's charity,
so that the afflicted and needy who have known your kindness
may one day receive  you thankfully
into the eternal dwelling of God.}\\

\Assembly{Amen.}\\

\Celebrant{And may the almighty God bless all of you, who are gathered here,
the Father, and the Son, \cross{} and the Holy Spirit.}\\

\Assembly{Amen.}\\

\Celebrant{In the Church and in the world may you be witnesses of the
gift of life and love that you celebrated. Go in peace.}\\

\Assembly{Thanks be to God.}

\newpage

\onelineheader{F}{inal song}{0mm}

\titleofsong{In eterno canter\`o (I will always sing to you)}

\textofsong{\songG}

\newpage

% restituisce il controllo al main, che lo restituisce al file 7 (retro di copertina, quasi uguale per la versione italiana e per quella inglese)
%%%
% gives back the control to the main file, that gives it to file 7 (back cover, almost the same for the Italian and for the English version)